Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2015-08-23 doc/GMP2015/revision_3/old_algo_sections.tex
Lots of work on Revision 3, getting close.

\section{Main Result: Canonical DR-Plan, Optimality, and Algorithm}
\label{sec:DRP}
% In the problem of the optimal DR-Plan there is generally not a unique plan.
% % Indeed, we will prove a union of $N$ isostatic subgraphs will result in $N$ unique plans, but that at the $N^{\text{th}}$ level of the tree it will always be the same. Therefore, all choices of decomposition are in some sense equivalent. The theorem we seek to prove is thus:
% However, we will show that regardless of which children are chosen for the plan, so long as they satisfy the definition of an optimal DR-plan, the recombination will require solving of the same systems. Being the smallest such structure that offers this, the definition of an optimal DR-plan could be considered the canonical DR-plan.
% % To assist in showing this, we prove this core theorem throughout this section:

% In this section, we discuss 2D bar-joint graphs. All vertex weights are $2$, all edge weights are $1$, and constant $k= -{{3}\choose{2}}=-3$. Trivial graphs are a single vertex and empty set. Furthermore, 2D isostatic graphs must be connected.
% % The greatest density of a 2D isostatic graph is $-2$ (the vertex). The other disconnected part of the graph would need to have a density of $-1$, which is overconstrained and not possible in a isostatic graph (because there is no trivial graph with that density).

\subsection{Canonical DR-Plan}
In this section, we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic the  uniqueness of a complete DR-plan, and we show that the nonunique aspects do not affect optimality for independent (underconstrained or isostatic) graphs. Furthermore, we give an efficient \ComplexityCanDRP\ algorithm to find the canonical DR-plan of any independent graph. The definition is as follows:

% \begin{definition}\label{def:canonical_drplan}
%     The \dfn{canonical DR-plan} of a graph $G$ satisfies the following three properties:
%     (1) it is a DR-plan of $G$;
%     (2) children are rigid vertex-maximal proper subgraphs of the parent; and
%     (3) if all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise exactly two that intersect non-trivially are children.
% \end{definition}

\begin{definition}\label{def:canonical_drplan}
    A \dfn{canonical DR-plan} is a DR-plan that satisfies the additional two properties:
    \begin{enumerate}
        \item Children are rigid vertex-maximal proper subgraphs of the parent.
        \item If all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise exactly two that intersect non-trivially are children.
    \end{enumerate}
\end{definition}

In this section and in section~\ref{sec:recomb}, any reference to a graph $G$ is assumed to be isostatic (i.e.\ well-constrained or $(k,l)$-tight).
% \sidenote{In this section, $G$ is always assumed to be isostatic.}

% This definition gives the canonical DR-plan a surprisingly strong Church-Rosser property.

Definition~\ref{def:canonical_drplan} gives the canonical DR-plan a surprisingly strong Church-Rosser property, which is made explicit in Theorem~\ref{theorem:main}, the main result of this section.

\begin{theorem}
\label{theorem:canonical_exists_and_is_optimal}
\label{theorem:canonical_is_optimal}
\label{theorem:main}
    A canonical DR-plan exists for a graph $G$ and any canonical DR-plan is optimal if $G$ is independent.
\end{theorem}

% \begin{theorem} \label{theorem:canonical_is_optimal}
%     \label{theorem:main}
%     Any canonical DR-plan is an optimal DR-plan.
% \end{theorem}

%The proof of this theorem is a direct consequence of the following  more general theorem.

%\begin{theorem}\label{theorem:main}
%Given an isostatic 2D bar-joint graph $G$ and ComDRP$(G)$, for all nodes $C$
%with children $C_1,\ldots,C_N$ preserve children according to the following rules.
%\begin{enumerate}
%    \item If $C_i \cap C_j$ is trivial then keep all $C_1,\ldots,C_N$ as children.
%    \item If $C_i \cap C_j$ is isostatic then select any two out of $C_1,\ldots,C_N$ as children.
%\end{enumerate}
%This is a canonical DR-plan.
%\end{theorem}

\begin{proof}
We show the existence of a canonical DR-plan by constructing it as follows:

Begin with $\comdrp{G}$ of a rigid 2D bar-joint graph $G$, for all nodes $C$ with children $C_1,\ldots,C_N$  retain children nodes according to the following rules:
\begin{enumerate}
   \item If $C_i \cap C_j$ is trivial then retain all $C_1,\ldots,C_N$ as children.
   \item If $C_i \cap C_j$ is rigid then select any two out of $C_1,\ldots,C_N$ as children.
\end{enumerate}

This directly satisfies Properties (2) and (3) of a canonical DR-plan (see Definition~\ref{def:canonical_drplan}), because all the nodes in $\comdrp{G}$ are rigid vertex-maximal proper subgraphs,  which we shorten to {\em clusters}.  To show Property (1) holds (that this constitutes a DR-plan):
for Case 1 above,  since we start with a complete DR-plan, if we preserve all the children it is still a DR-plan; for Case 2 above, we know that the union must be rigid as well and it cannot be anything other than $C$, otherwise we would have found a larger rigid proper subgraph of $C$, contradicting vertex-maximality.

Note that if we begin with an isostatic graph, ``rigid'' can be replaced with ``isostatic'' throughout the construction and preserve the above properties. The rigid proper subgraphs of an isostatic graph must be isostatic themselves.

Next we show that a canonical DR-plan is optimal.

First, note  that any  DR-plan $R$ without the Property (2) of a canonical DR-plan can always be modified (by introducing intermediate nodes) to satisfy Property (2) without  increasing the max fan-in, since any rigid proper subgraph of a graph $C$ (a child of node $C$ of the DR-plan $R$)
is the subgraph of some cluster of $C$.
Thus without loss of generality, we can assume that an optimal DR-plan satisfies Property (2) of a canonical DR-plan.

The proof of optimality of a canonical DR-plan is by induction on its height.  The base case trivially holds for canonical DR-plans of height 0, i.e.\ for single edges. The induction hypothesis is that canonical DR-plans of height $t$ are optimal for the root node.
For the induction step consider a canonical DR-plan $R$ of  height $t+1$ rooted at a node $C$.
Notice that $R$ represents a canonical DR-plan $R(C)$ for the graphs $C$ corresponding to each of its  descendant nodes.
Thus, from the induction hypothesis, we know that the $R(C_i)$ is optimal for $C_i$.

Thus it is sufficient to demonstrate a set of nodes $S$  that must be present in any  DR-plan $R$ for $C$ that satisfies Property (2), including a known optimal one; and furthermore, for any such DR-plan $R$,  either (Claim 1) $S$ must be the set of children of $C$; or (Claim 2) for all the ancestors $A$ of $S$, $R$ has the minimum possible fan-in of 2.

We show the two claims below.
The first claim is that for a node $C$ whose clusters have trivial pairwise intersections, any DR-plan of $C$ that satisfies Property (2) must also satisfy Property (3) at $C$, i.e.\ the set of children $S$ of $C$ consists of all clusters of $C$.
Because this is the only choice, it is the minimum fan-in at $C$ for any DR-plan for $C$ with Property (2), including a known optimal one.
The second claim shows that in the case of nodes $C$ whose rigid, vertex-maximal proper subgraphs have  non-trivial pairwise intersections, every canonical DR-plan of $C$ that uses any possible choice of two such subgraphs of $C$ as children results in a minimum possible fan-in of 2 in the ancestor nodes $A$ leading to the {\em same maximal antichain $S$ of descendants $D$ of $C$}. The antichain is maximal in the partial order of rigid subgraphs of $C$ under containment. I.e.\ $S$ satisfies the property that every proper vertex-maximal rigid subgraph of $C$ is a superset of some $D$ in $S$; this follows from properties of maximal antichains that no element of $S$ is contained in the union of other elements of $S$; and the union of elements of $S$ is $C$. Thus any  DR-plan that satisfies Property (2) and hence contains two or more of the rigid vertex-maximal proper subgraphs of $C$ as children must also contain every element of $S$. The two claims complete the proof that every canonical DR-plan is optimal.

 % We do this inductively on the l, beginning with the root node, showing that the rules are always the optimal choice.
% by discussing each rule and proving inductively that this is the optimal choice at each level of the DR-plan tree, starting with the root node.

% Now we show that this plan is optimal by considering the two cases. Observation \ref{lemma:union_intersection} shows that we do not have to consider anything other than the two cases stated in the construction.
 % the intersection of any two isostatic subgraphs can only result in trivial or isostatic subgraphs. Therefore, given $C$ and its isostatic vertex-maximal subgraphs $C_1,\ldots,C_N$, the are only two possibilities to consider. Either (1) subgraphs $C_i$ and $C_j$ have a trivial intersection, or (2) they have an isostatic intersection.

\medskip\noindent
\vemph{Claim 1:}
% If some pair is intersecting trivially, then in fact all pairs intersect trivially, thus all must be children by defn of canonical drplan
%
A set of clusters $C_1,\ldots,C_N$ whose pairwise intersection is trivial, must be children of $C$ in an optimal DR-plan.

We prove this claim  by showing that the union of no subset of the children can be $C$, thereby requiring all of them to be included as children.

% must be optimal for this node.

We prove by contradiction.
Assume to the contrary that the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ is isostatic. If $U\neq C$, then we found a larger proper subgraph contradicting vertex-maximality of the $C_i$. So, it must be that $U=C$.
\usestwod
However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:combined_lemma}, Item \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial, i.e.\ disconnected vertices. By definition of a DR-plan, $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$. Thus, $C_k$ is (i) a collection of disconnected vertices, and (ii) an isostatic subgraph of $C$, which is impossible. As $C$ is isostatic, this means the union of no proper subset of $C_1,\ldots,C_N$ is isostatic, nor is it equal to $C$, proving Claim 1.

Furthermore, since a canonical DR-plan has nodes with proper rigid \vemph{vertex-maximal} subgraphs as children, if, as in this case, their pairwise intersection is trivial, it follows that any node has at most as many children as a DR-plan without this restriction, because the union of the children must contain all edges of the parent. Therefore, the canonical DR-plan is the optimal choice in this case of trivial intersections.

\medskip\noindent
\vemph{Claim 2:}
If some pair in the set of child clusters $C_1,\ldots,C_N$ of $C$ has an isostatic (nontrivial) intersection, then choosing any two as children (minimum possible fan-in) will result in the same maximal antichain of descendants of $C$.

To prove Claim 2, notice that if  $C_i \cap C_j$ is isostatic, then, by Observation \ref{lemma:union_intersection}, $C_i \cup C_j$ is also isostatic. This means that, by Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc}, the union of any two children of $C$ is $C$ itself. Thus, any two children can be chosen to make a canonical DR-plan and that is the minimum fan-in possible for a node of the DR-plan.
% are potential choices for the optimal DR-plan as they all create equal fan-in (exactly two) at this level.

\newcommand{\induceonc}[1]{Idc\left(C,#1\right)}
\renewcommand{\induceonc}[1]{#1}
\newcommand{\iunion}[1]{\induceonc{I\cup\bigcup_{k\in [N]\setminus\{#1\}}{R_k}}}

However, to guarantee that any two are the \vemph{optimal} choice, it must ensure minimum fan-in over all descendants leading up to a common maximal antichain $S$ of subgraphs.

To prove this holds, take the set $[N]=\{1,\dots,N\}$, and denote $I=\bigcap_{k\in [N]}{C_k}$ and $R_k=C\setminus C_k$. Suppose  $C_i$ and $C_j$, where $i\neq j$,  are  the children. For convenience, we will assume all subgraphs are induced subgraphs of $C$. We know that $C=\induceonc{I\cup\bigcup_{k\in [N]}{R_k}}$ and $C_i=\iunion{i}$. The isostatic vertex-maximal subgraphs of $C_i$ are $(\iunion{i,1}),\ldots,(\iunion{i,i-1}),(\iunion{i,i+1}),\ldots,(\iunion{i,N})$ all of whose pairwise intersections are isostatic subgraphs. So any two of these are viable children for $C_i$.
% Since
% \[C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)\]
% the children of $C_i$ will be
% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)\]
% and
% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)\]
% % $C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)$
% % the children of this node will be
% % $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)$
% % and
% % $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)$
% for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other. \todo{Prove these are valid children? Or is this obvious?}
This continues for $N-1$ levels, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $\induceonc{I\cup R_k}$ for $k\in [N]$, with every $k$ appearing at least once. At the last level, there are exactly two rigid proper vertex-maximal subgraphs, and hence a unique choice of pair of children. Thus, regardless of the sequence of choices of $C_i$ and $C_j$, and of their descendants at each level, the DR-plan has the optimal fan-in of two for every node for  $N$ levels,  and the collection of last level nodes contain the same maximal antichain of subgraphs (for all choices).
%
% \medskip\noindent
% This proof then applies to itself recursively to show that the fan-in of the children will also be minimum.
%
% -- say this at top `proof is by induction on the level of the dr=plan'
\end{proof}

This proof of this theorem relies on the following crucial observation and lemma. These will be used again in the application sections (\ref{sec:bodypin} and \ref{sec:pinnedline}) of the paper, with modifications to work for other types of qusecs.

\begin{observation*}\label{lemma:union_intersection}
If $F_i$ and $F_j$ are non-empty isostatic graphs then the following hold: \\
(1) $F_i\cup F_j$ is not trivial;
(2) $F_i\cup F_j$ is underconstrained if and only if $F_i\cap F_j$ is trivial;
(3) $F_i\cup F_j$ is isostatic if and only if $F_i\cap F_j$ is isostatic; and
(4) $F_i\cap F_j$ is not underconstrained.
\end{observation*}

The following key properties hold at the nodes of a canonical DR-plan.

\begin{lemma*}\label{lemma:combined_lemma}
Let $C$ be an isostatic node of a canonical DR-plan, with distinct children $C_1,C_2,\ldots, C_m$. Assume $i\ne j$.
Then
\begin{enumerate}
    \item\label{lemma:wc_intersection_is_C}
    $C_i\cup C_j$ is isostatic if and only if $C_i\cup C_j = C$.

    \item\label{lemma:wc_intersection_makes_all_wc}
    If $C_i\cup C_j$ is isostatic, then $\forall k: C_i\cup C_k$ is isostatic. Alternatively, if $C_i\cup C_j=C$, then $\forall k: C_i\cup C_k=C$.

    \item\label{lemma:uc_intersection_makes_all_uc}
    If $C_i\cap C_j$ is trivial, then $\forall k: C_i\cap C_k$ is trivial.
\end{enumerate}
\end{lemma*}

\begin{remark}
The first item in the above lemma generalizes to the union of any number of children, $C_1,\ldots,C_k$, resulting in the desirable property of \dfn{cluster minimality} (defined in \cite{hoffman2001decompositionI} and in Section \ref{sec:prev}) holding for canonical-optimal DR-plans.
\end{remark}

% \begin{figure*}\centering
% \begin{subfigure}{.3\linewidth}\centering
%     \input{../../img/tikz/c2c3_of_k33s-0.tikz}
%     \caption{}\label{fig:c2c3ofk33s:a}
% \end{subfigure}%
% \begin{subfigure}{.7\linewidth}\centering
%     \input{../../img/tikz/c2c3_of_k33s-1.tikz}
%     \caption{}\label{fig:c2c3ofk33s:b}
% \end{subfigure}

% \caption{(\ref{fig:c2c3ofk33s:a}) A doublet ($C_2 \times C_3$) with each edge of the triangles replaced by a $K_{3,3}$. This pattern continues inwards for a total of $N$ triangles, indicated by the dashed lines. (\ref{fig:c2c3ofk33s:b}) Most of the DR-plan of this graph, omitting further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges and of edges into the component nodes. $G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs. The dashed lines indicated that this exact structure is repeated.}
% \label{fig:c2c3ofk33s}
% \end{figure*}

% \FigInit
%     {(\ref{fig:c2c3ofk33s:a}) A doublet ($C_2 \times C_3$) with each edge of the triangles replaced by a $K_{3,3}$. This pattern continues inwards for a total of $N$ triangles, indicated by the dashed lines. (\ref{fig:c2c3ofk33s:b}) Most of the DR-plan of this graph, omitting further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges and of edges into the component nodes. $G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs. The dashed lines indicated that this exact structure is repeated.}
%     {fig:c2c3ofk33s}
% \FigTwoSubfigWithWidth
%     {.3}
%     {../../img/epsfromtikz/c2c3_of_k33s-0}
%     {}
%     {fig:c2c3ofk33s:a}
%     %
%     {.7}
%     {../../img/epsfromtikz/c2c3_of_k33s-1}
%     {}
%     {fig:c2c3ofk33s:b}

% \ClearMyMinHeight
% \SetMyMinHeight{.3}{../../img/epsfromtikz/c2c3_of_k33s-0}
% \SetMyMinHeight{.7}{../../img/epsfromtikz/c2c3_of_k33s-1}

% \begin{figure*}\centering%
%     %
%     \begin{subfigure}{0.3\linewidth}\centering
%         \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/c2c3_of_k33s-0}
%         \caption{}\label{fig:c2c3ofk33s:a}
%     \end{subfigure}%
%     %
%     \hfill
%     \begin{subfigure}{0.7\linewidth}\centering
%         \includegraphics[height=\myMinHeight]{../../img/epsfromtikz/c2c3_of_k33s-1}
%         \caption{}\label{fig:c2c3ofk33s:b}
%     \end{subfigure}%
%     %
%     \caption{(\ref{fig:c2c3ofk33s:a}) A doublet ($C_2 \times C_3$) with each edge of the triangles replaced by a $K_{3,3}$. This pattern continues inwards for a total of $N$ triangles, indicated by the dashed lines. (\ref{fig:c2c3ofk33s:b}) Most of the DR-plan of this graph, omitting further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges and of edges into the component nodes. $G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs. The dashed lines indicated that this exact structure is repeated.}
%     \label{fig:c2c3ofk33s}
% \end{figure*}%

\ClearMyMinHeight
\SetMyMinHeight{.3}{../../img/svg/revised_c2c3_of_k33s}
\SetMyMinHeight{.7}{../../img/svg/revised_c2c3_of_k33s_candrp}

\begin{figure*}\centering%
    %
    \begin{subfigure}{0.3\linewidth}\centering
        \includegraphics[height=\myMinHeight]{../../img/svg/revised_c2c3_of_k33s}
        \caption{}\label{fig:c2c3ofk33s:a}
    \end{subfigure}%
    %
    \hfill
    \begin{subfigure}{0.7\linewidth}\centering
        \includegraphics[height=\myMinHeight]{../../img/svg/revised_c2c3_of_k33s_candrp}
        \caption{}\label{fig:c2c3ofk33s:b}
    \end{subfigure}%
    %
    \caption{
    (\ref{fig:c2c3ofk33s:a}) A sequence of doublets ($C_2 \times C_3$) intersecting on triangles, where the edges of the triangles are replaced by $K_{3,3}$'s. This pattern continues inwards for a total of $N$ triangles, indicated by the dashed arrows. (\ref{fig:c2c3ofk33s:b}) The canonical DR-plan of $G$, drawn as a DAG. $G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs. Below the third level, the obvious pattern continues until only the individual doublets are present (fourth level) with the ellipses indicating the remaining doublets between those shown. Decomposition of one of these doublets is shown. The dashed lines indicated that this exact decomposition (of the similar nodes on the level) is repeated. Further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges is omitted from the figure.
    }
    \label{fig:c2c3ofk33s}
\end{figure*}%

\begin{example}[DR-plan for self-similar structure]
% \myexample
% \textsl{[DR-plan for self-similar structure]}
This example details the decomposition of the graph in Figure \ref{fig:c2c3ofk33s}, the canonical DR-plan of $G$. It begins with the whole (isostatic) graph as the root. The graph $G$ has only two isostatic vertex-maximal subgraphs: $G$ without the outermost triangle composed of $K_{3,3}$ graphs (triangle $1$) and $G$ without the inner triangle (triangle $N$). These intersect on $G$ without triangle $1$ and $N$ which is clearly isostatic. As explained in the proof of Theorem \ref{theorem:main}, since there are only 2 possible children, their intersection must be a node 2 levels below the parent. As expected, it is on the third level, as a child of both of $G$'s children.

Both of $G$'s children are similar to $G$, but containing only $N-1$ triangles. Therefore, the canonical DR-plans of these children follow the same pattern. This continues downward until the individual doublets are reached (there will be multiple occurrences of the same doublets at this level, but they can be represented as the same node in a DAG).

Further decomposition of one of these doublets is shown. The three edges between the triangles and the triangles themselves all intersect trivially pairwise. By Theorem \ref{theorem:main}, part 1, they must all be children in the DR-plan. Similarly, the triangles decompose into their three trivially intersecting $K_{3,3}$'s. Then the $K_{3,3}$ subgraphs decompose into their separate 9 edges.

The self-similar nature of this graph is evident in the canonical DR-plan. Many structures are repeated throughout the DR-plan, allowing for shared computation in both decomposition and recombination.
\end{example}

% \subsection{Extensions}
% This framework immediately pushes through for body-pin systems via a simple reduction. If there are $N$ pins on a body, it can be represented as a 2-tree with $N$ vertices, each corresponding to a pin, making sure to select edge distances such that the distance between pins is preserved. E.g.\ a body with two pins is an edge, three pins is a triangle, etc. Any bodies that share a pin now intersect on their vertex that corresponds to that pin. Now we have a bar-joint representation of the body-pin system in 2D and all proofs follow.

% With more effort, it can be shown that pinned line-incidence systems can also use this framework. This is done in section \ref{XXX}.

\subsection{Algorithm}

@@

\begin{theorem}\label{theorem:algo_complexity}
    There exists an \ComplexityCanDRPV\ algorithm to find a canonical DR-plan for an independent input graph.
\end{theorem}

@@

% In the case of trivial intersections, all of the found clusters will be unique. If the cardinality is greater than 2, we can save computation by not checking the other edges in the found cluster. But we can only know that they are trivial intersections after finding 2 different clusters. So do that first...

% In the case of isostatic intersections, if we begin with an edge in the core, there are several clusters containing this edge (and the algo will pick the largest among them). We could only safely skip the other edges in the core, but we don't know what it is yet. So after finding the first cluster, go to an edge not in it and find the next cluster. Check the intersection. Now we can skip the others in the core. WAIT.... THIS WILL ONLY PICK THE LARGEST, AND ITS ``COMPLEMENT'', SO TO SPEAK. WHICH TECHNICALLY WORKS....

% For each found cluster $c$, for each edge $e$ \emph{not in} $c$, delete $c$ from $M(e)$ and delete the pointer to $M(e)$ from the pointer list of $e$.

@@

% \begin{figure*}\centering%
%   \includegraphics[width=\linewidth]{../../img/svg/overconstrained_choices}
%   \caption{The first graph is over constrained by one edge (a $K_{3,3}$ with an extra vertex attached by 3 edges). The second graph is a minimally rigid subgraph of the first. Any DR-plan of this graph will have fan-in of at least 9, occurring at the node that decomposes the $K_{3,3}$. The third graph is a different subgraph of the first. A canonical and cluster minimal DR-plan of size 2 can be made for this graph (it is a Henneberg-I construction). Thus, we can see the first graph has both non-optimal (second graph) and optimal (third graph) canonical, cluster minimal DR-plans.}
%   \label{fig:overconstrained_choices}
% \end{figure*}%

\ClearMyMinHeight
\SetMyMinHeight{.6}{../../img/svg/new_overconstrained_not_optimal}
\SetMyMinHeight{.4}{../../img/svg/new_overconstrained_optimal}

\begin{figure*}\centering%
    %
    \begin{subfigure}{0.5\linewidth}\centering
        \scalebox{1}[.9]{
        \includegraphics[height=\myMinHeight]{../../img/svg/new_overconstrained_optimal}
        }
        \caption{}\label{fig:overconstrained:optimal}
    \end{subfigure}%
    %
    \hfill
    \begin{subfigure}{0.5\linewidth}\centering
        \scalebox{1}[.9]{
        \includegraphics[height=\myMinHeight]{../../img/svg/new_overconstrained_not_optimal}
        }
        \caption{}\label{fig:overconstrained:not_optimal}
    \end{subfigure}%
    %
    \caption{
    % (\ref{fig:overconstrained:optimal}) An optimal DR-plan of a singly overconstrained rigid graph with a fan-in of 5. Decomposition of triangles is omitted. (\ref{fig:overconstrained:not_optimal}) A canonical and cluster-minimal DR-plan of the same graph. The DR-plan has a fan-in of 9 and is non-optimal, shown by the preceding counter-example. Since the nontrivial intersection of the two children of the root is underconstrained, their decompositions are separate, causing the large fan-in.
    Both figures are canonical and cluster-minimal DR-plans of the same singly overconstrained rigid graph. Decomposition of triangles are omitted and dashed lines indicate a decomposition similar to the other nodes on the same level. (\ref{fig:overconstrained:optimal}) is an optimal DR-plan, with a fan-in of 5. (\ref{fig:overconstrained:not_optimal}) has a fan-in of 9 and is non-optimal, shown by the preceding counter-example.
    }
    \label{fig:overconstrained}
\end{figure*}%

\subsection{Overconstrained Graphs and NP-Hardness of Optimal DR-Planning}
\label{sec:drp:overconstrained}

For overconstrained (not independent) graphs, a canonical DR-plan is still well-defined.
However, it may be far from optimal. The proofs of Theorem \ref{theorem:main}, Observation \ref{lemma:union_intersection}, and Lemma \ref{lemma:combined_lemma} all fail for overconstrained graphs.
It is important to note that, regardless whether the graph is overconstrained, if every node in a canonical DR-plan $R$ has clusters whose pairwise intersection is trivial, then the DR-plan is the unique one satisfying Property (2), and since we know that there is an optimal DR-plan that satisfies Property (2), $R$ is in fact optimal. The problem arises when some node in a DR-plan has clusters whose pairwise intersection is non-trivial.
In this case, an arbitrary choice of a pair of clusters as children of an overconstrained node in a canonical DR-plan may not result in an optimal DR-plan. This is in contrast to independent graphs, which, as shown in Theorem~\ref{theorem:main}, exhibit the strong Church-Rosser property that any choice yields an optimal DR-plan.
% \todo{rephrase? Theorem 4 shows this. If it's independent it \vemph{will} be optimal (the church rosser property), no longer holds if it's overconstrained.}
A good source of examples of overconstrained graphs with canonical DR-plans that are not optimal are graphs whose cluster-minimal DR-plans that are not optimal. The example shown in Figure \ref{fig:overconstrained} is a canonical, cluster-minimal DR-plan that is not optimal; an optimal DR-plan is also shown in the figure.
The root cause of the NP-hardness is encapsulated in this figure: because the different choices of vertex-maximal subgraphs for overconstrained input do not incur the same fan-in, finding the optimal DR-plan becomes a search problem with a combinatorial explosion of options.

As mentioned earlier, the Modified \frontier\ algorithm
version given in \cite{lomonosov2004graph} runs in polynomial time and finds a cluster-minimal DR-plan for any graph.
Similarly, the algorithm given above finds a canonical DR-plan also for any input graph.  However neither of these DR-plans may be optimal for overconstrained graphs as shown in Figure \ref{fig:overconstrained}.

While the canonical DR-plan is optimal only if the input graph is independent, when there are only $k$ overconstraints for some fixed $k$, we can still find the optimal DR-plan using a straightforward modification of the above algorithm. However, the time complexity is exponential in $k$.

This exponential growth of time complexity for overconstrained graphs is in fact captured in the proof of NP-hardness of optimal DR-planning in \cite{sitharam2005combinatorial, lomonosov2004graph}.

>>>>

@@

\begin{theorem}
\label{theorem:algo_complexity}
    There exists an \ComplexityCanDRPV\ algorithm to find a canonical DR-plan for an independent input graph.
\end{theorem}

@@

@@

While the canonical DR-plan is clear conceptually, it is difficult to reason with directly in the design and analysis of algorithms. Therefore, we introduce the following modification to get a similar but more useful structure.

\begin{definition}
    A \dfn{sequential} canonical DR-plan (or simply sequential DR-plan) is a canonical DR-plan with the following modification:

    In the case that a node has isostatic vertex-maximal proper subgraphs (or clusters) that intersect non-trivially.
    we construct a \dfn{sequential extension} in the following manner.
    %
    Let the current node be $G$.
    Let the $N$ clusters be $C_1, C_2, \ldots, C_N$.
    Let the \dfn{core} be $I=\bigcap_{k=1}^{N}{C_k}$.
    Let the $N$ \dfn{appendages} be $R_i = G\setminus C_i$.
    %
    Choose arbitrary appendage $R_i$ and set the children of $G$ as $I\cup \bigcup_{k=1}^{i}{R_k}\cup \bigcup_{k=i}^{N}{R_k}$ and the clusters of $R_i$. Construct the sequential canonical DR-plan of each child.
\end{definition}

See Figure~\ref{fig:demo_graph:candrpseq} for an example. Note the pattern will continue for $N$ levels, where $N$ is the number of appendages. At the top is $G$, the core with all appendages; at the next level, the children are the core with all but one appendage and the clusters of the removed appendage; at the next level, it has the core with all but two appendages and the clusters of the appendage removed second; etc.; at the bottom level is the core.

\begin{observation}
    A sequential canonical DR-plan of independent graph $G=(V,E)$ is a tree with $O(|V|)$ nodes.
\end{observation}

\begin{observation}
    A sequential canonical DR-plan of an independent graph is optimal.
\end{observation}

\begin{proof}
    A sequential DR-plan behaves the same as a canonical DR-plan in the case of a node whose clusters have trivial intersections.
    In a regular canonical DR-plan, at a node with $N$ clusters that intersect non-trivially, there is fan-in of 2 until level $N$. At level $N-1$, there are $N$ distinct nodes, one for each appendage together with the core. The children of these nodes are the core and the clusters of the corresponding appendage. Thus, the maximum fan-in over this part of the canonical DR-plan is one plus the maximum number of clusters in an appendage. By construction, this is exactly the maximum fan-in of a sequential extension.
\end{proof}

\begin{observation}\label{obs:seq-to-reg-canonical}
    Given a sequential canonical DR-plan of an independent graph $G=(V,E)$, a canonical DR-plan can be found in time $O(|V|^3)$.
\end{observation}

\begin{proof}
    In the case of a node whose children intersect trivially, nothing needs to be done.
    In the case of non-trivial intersections, we will need to construct the canonical DR-plan subtree from the sequential extension.

    At such a node $G$ with $N$ clusters, we will have to build $\sum_{i=1}^N{i}=O(N^2)$ new nodes. For convenience, $C_{[i,j]}$ will represent the core plus appendages $i$ through $j$. We set the children of $G$ as $C_{[0,N-1]}$ and $C_{[1,N]}$. The children of $C_{[0,N-1]}$ are set as $C_{[0,N-2]}$ and $C_{[1,N-1]}$. The children of $C_{[1,N]}$ are set as $C_{[1,N-1]}$ (not unique) and $C_{[2,N]}$. This pattern continues for $N$ levels, with level $i$ containing $i$ unique nodes.

    Since there are at most $O(|V|)$ nodes with $O(|V|)$ clusters that need this treatment, the complexity is $O(|V|^3)$.
\end{proof}

Before discussing the algorithm, we remind the reader of the component pebble game~\cite{Jacobs:1997:PG}. Given an independent input graph $G$ with vertex weight 2 and edge weight 1, the output will be all of the 2D maximal isostatic subgraphs (the components). We will denote this algorithm with $\components{G}$. For independent input, this runs in time $O(|V|^2)$.

The algorithm for finding the sequential DR-plan is going to rely on the observation that the set of $\components{G\setminus e}$ over all $e\in E$ will (essentially) encompass all the nodes of the sequential DR-plan for $G$, with minor caveats that we will explain later.

\begin{lemma}\label{lemma:seq-nontriv-Gtoe}
    When the intersections of clusters are trivial at all nodes in the (sequential) canonical DR-plan, then the set $\components{G\setminus e}$ will contain all of the siblings in the path from $G$ to $e$.
\end{lemma}

\begin{proof}
    The sequential and canonical DR-plans are identical in this case. Suppose $G$ has $N$ trivially intersecting components, and $e$ is in component $C_i$. Removing edge $e$ will leave all components besides $C_i$ intact, thus they will be isostatic components in $G\setminus e$ and will be in the set. Therefore, $\components{G\setminus e}$ contains
    % $C_1, \ldots, C_{i-1}, C_{i+1}, \ldots, C_N$
    all clusters besides $C_i$
    and the set $\components{C_i\setminus e}$. By induction, the proof is complete.
\end{proof}

\begin{lemma}\label{lemma:seq-Gtoe}
    In the (sequential) canonical DR-plan, the set $\components{G\setminus e}$ will contain all of the siblings in the path from $G$ to $e$ excluding the following: in the case of a node whose $N$ clusters have non-trivial intersections, those nodes that contain the core plus any $i$ appendages where $0\leq i<N-1$.
\end{lemma}

\begin{proof}
    By Lemma~\ref{lemma:seq-nontriv-Gtoe}, this is true for trivial intersections. For non-trivial intersections, consider an edge $e$ that comes from appendage $R_i$. Then $\components{G\setminus e}$ will contain the element `core plus all appendages but $R_i$' and the set $\components{R_i\setminus e}$, so the statement is true. Now, consider an edge $e$ that comes from $I$, the core. Then $\components{G\setminus e}$ will contain $\components{R_i}$ for all appendages $R_i$ and $\components{I\setminus e}$.
\end{proof}

Note that in the sequential DR-plan, the number of missing nodes is at worst $O(|V|^2)$. There can be as many as $O(|V|)$ nodes with non-trivial intersections on $O(|V|)$ clusters, each cluster contributing one appendage and one missing node (excluding the first in the sequential extension).

\begin{theorem}\label{theorem:algo_complexity}
    There exists an \ComplexityCanDRPV\ *** algorithm to find a canonical DR-plan for an independent graph.
\end{theorem}
